---
title: "poisson"
author: "Gulzina Kuttubekova"
date: "5/13/2020"
output: pdf_document
---


```{r, include=FALSE}
library(ggplot2)
library(ggpubr)
library(dplyr)
library(reshape)
library(LearnBayes)
library(rmutil)
```

Read the dataset:
```{r}
covid <- read.csv("~/Desktop/spring2020/stat207/bayesian-analysis-of-COVID19-in-CA/data/Covid19-04-13-20.txt")

# scale population by 1000
covid$Population <- covid$Population/1000
```


Sample from poisson using two-step direct sampling. 
```{r}
#' Draw one sample from full conditional lambda distribution for all counties
#' 
#' @param alpha one sample of alpha
#' @param beta one sample of beta
#' @param C population of 58 counties
#' @param N number of infected people in 58 counties 
#' @return 1*58-dim vector with samples of lambda for all counties
conditional_lambda <- function(alpha, beta, C, N) {
    s <- c()
    
    for (i in 1:58) {
        alpha_star <- alpha + N[i]
        beta_star <- 1 / (1/beta + C[i]/1000)
        
        s[i] <- rgamma(1, shape = alpha_star, scale = beta_star)
    }
    
    return(s)
}
```

```{r}
#' Evaluate the ratio of two gamma functions using Stirling's formula
#' 
#' @param alpha
#' @param n
#' @return
gamma_ratios <- function(alpha, n) {
    res <- ((n + alpha - 1) / (n - 1))^n
    res <- res*((n + alpha - 1) / exp(1))^alpha
    
    return(res)
}
```


```{r}
#' Evaluate p(alpha, beta|n)
#' 
#' @param alpha parameter
#' @param beta parameter
#' @param C population of 58 counties
#' @param N number of infected people in 58 counties
#' @return value of marginal evaluated at alpha, beta 
marginal_alpha_beta <- function(alpha, beta, C, N) {
    pv <- c()
    # scale N by 100
    N <- N
    
    for (i in 1:58) {
        c = C[i]
        n = N[i]
        
        if (n == 0) {
            pv[i] = 1 / ( (beta*(c + 1/beta))^alpha )
        } else if (n ==1) {
            num = c*gamma(alpha+1)
            denom = beta^alpha * gamma(alpha) * (c + 1/beta)^(alpha+1)
            pv[i] = num / denom
        } else {
            gamma_ratio = gamma_ratios(alpha, n)
            powers = (c*beta/(c*beta + 1))^n * (beta/(c*beta + 1))^beta
        
            num = gamma_ratio * powers
            denom = n * beta^(alpha) * gamma(alpha)
            pv[i] <-  10^25 * num / denom
        }
    }
    
    res <- dgamma(alpha, 
                  shape = 10, 
                  scale = 2) * dgamma(beta, 
                                      shape = 20,
                                      scale = 0.5) * prod(pv)
    
    return(res)
}
```




Sample 500 samples from joint marginal of alpha and beta:
```{r}
# Grid to evaluate in dataframe
dt <- expand.grid(X = seq(0.1, 40, length=500), 
                  Y = seq(0.1, 30, length=500))

# get probabilities
for (i in 1:nrow(dt)) {
    dt$p[i] <- marginal_alpha_beta(dt[i,1], dt[i,2], 
                                   covid$Population,
                                   covid$Total.cases)
}

# sample from marginal 
n = 50000

# take only the ones with non inf values
dt <- dt[!is.infinite(rowSums(dt)),]

# normalize probs
dt$p <- dt$p / sum(dt$p)

# sample from it
idx <- sample(1:nrow(dt), size = n, replace=TRUE, prob=dt$p)
sampled_alpha <- dt$X[idx]
sampled_beta <- dt$Y[idx]
```

```{r}
sampled_alpha %>% hist()
sampled_beta %>% hist()
```
Posterior estimates
```{r}
quantile(sampled_alpha, probs = c(0.025, 0.5, 0.975))
quantile(sampled_beta, probs = c(0.025, 0.5, 0.975))
```


Posterior and prior distributions of alpha and betas:
```{r}
sa <- data.frame(posterior = sampled_alpha, 
                 prior = rgamma(n, 10, scale = 2))
sa_melt <- melt(sa, measure.vars = c('posterior', 'prior'))
colnames(sa_melt) <- c('type', 'alpha')
```

```{r}
pa <- ggdensity(sa_melt, 'alpha', fill = 'type') + 
    xlim(0, 50)
```

```{r}
sb <- data.frame(posterior = sampled_beta, 
                 prior = rgamma(n, 20, scale = 0.5))
sb_melt <- melt(sb, measure.vars = c('posterior', 'prior'))
colnames(sb_melt) <- c('type', 'beta')
```

```{r}
pb <- ggdensity(sb_melt, 'beta', fill = 'type') + 
    ylab("") + 
    xlim(0, 20)
```

```{r}
ggarrange(pa, pb, common.legend = T)
```


Sample lambdas:
```{r}
lambdas <- c()
for (i in 1:n) {
    temp <- conditional_lambda(sampled_alpha[i],
                               sampled_beta[i],
                               covid$Population,
                               covid$Total.cases)
    lambdas <- rbind(lambdas, temp)
}
```

```{r}
lambdas %>% dim()
```

```{r}
lambdas %>% apply(2, quantile, probs = c(0.025, 0.5, 0.975))
```

Plot all lambdas as boxplots:
```{r}
# rename columns
colnames(lambdas) <- covid$County

# melt
lambdas_melt <- melt(lambdas)

# boxplots
lambdas_melt %>% ggplot(aes(y=value, x=X2)) + 
    geom_boxplot(outlier.size = 0.5) + 
    theme(axis.text.x = element_text(angle = 90, size = 6)) + 
    ggtitle('Distribution of lambda by county') +
    xlab("") + 
    ylab("lambda")
```

Estimate mean infections for two different counties:
```{r}
mean(lambdas[1]) * covid$Population[1]
mean(lambdas[7]) * covid$Population[7]
```


Obtain samples from posterior predictive distribution:
```{r}
#' Draw one sample from predictive distribution for all counties
#' 
#' @param lambda one sample of lambda
#' @param C population of 58 counties
#' @return 1*58-dim vector with samples of n for all counties
predictive_n<- function(lambda, C) {
    s <- c()
    
    for (i in 1:58) {
        s[i] <- rpois(1, lambda = lambda[i]*C[i])
    }
    
    return(s)
}
```

```{r}
n <- nrow(lambdas)
ns <- c()
for (i in 1:n) {
    temp <- predictive_n(lambdas[i,], covid$Population)
    ns <- rbind(ns, temp)
} 
```

Use this samples along with thetas from previous study:
```{r}
ds <- matrix(NA, nrow = nrow(ns), ncol = ncol(ns))

# estimate E[theta|y]
mean_thetas <- apply(thetas, 2, mean)

# multiply each mean theta by n_i's
for (i in 1:58) {
    ds[,i] <- ns[,i]*mean_thetas[i] / covid$Population[i] 
}
```

Boxplot for death per 1000 habitants distirbution
```{r}
# rename columns
colnames(ds) <- covid$County

# melt
ds_melt <- melt(ds)

# boxplots
ds_boxplots <- ds_melt %>%  
    ggplot(aes(y=value, x=X2)) + 
    geom_boxplot(outlier.size = 0.5) + 
    theme(axis.text.x = element_text(angle = 90, size = 6)) + 
    ggtitle('Distribution of number of deaths per 1000 habitants') +
    xlab("") + 
    ylab("deaths per 1000")
```

```{r}
ds_boxplots
```


Top five counties:
```{r}
ds %>% apply(2, mean) %>% sort()
```

```{r}
covid$Deaths / covid$Population
```



Posterior predictive checks:

```{r}
# to see if Los Angeles is captured
maxs <- ns %>% apply(1, max)
sum(maxs == ns[,1])
```

```{r}
# percentage captured by LA only
la_real <- covid$Total.cases[1] / sum(covid$Total.cases)

# sampled
la_share <- c()
for (i in 1:nrow(ns)) {
    la_share[i] <- ns[i,1] / sum(ns[i,])
}
```

```{r}
hist(la_share, main = 'aksd')
abline(v = la_real, col="red")
```

```{r}
# percentage captured by Orange only
orange_real <- covid$Total.cases[5] / sum(covid$Total.cases)

# sampled
orange_share <- c()
for (i in 1:nrow(ns)) {
    orange_share[i] <- ns[i,5] / sum(ns[i,])
}
```

```{r}
hist(orange_share)
abline(v = orange_real, col='red')
```

```{r}
# percentage captured by Orange only
sf_real <- covid$Total.cases[7] / sum(covid$Total.cases)

# sampled
sf_share <- c()
for (i in 1:nrow(ns)) {
    sf_share[i] <- ns[i,7] / sum(ns[i,])
}
```

```{r}
hist(sf_share)
abline(v = sf_real, col='red')
```

```{r}
par(mfrow = c(3,4))
n <- nrow(ns)

for (j in 1:12) {
    # percentage captured by Orange only
    real <- covid$Total.cases[j] / sum(covid$Total.cases)

    # sampled
    share <- c()
    for (i in 1:n) {
        share[i] <- ns[i,j] / sum(ns[i,])
    }
    
    # plot
    share %>% hist(ylab = "", main = covid$County[j])
    abline(v=real, col="red")
}
```

```{r}
par(mfrow = c(3,4))
n <- nrow(ns)

for (j in 13:24) {
    # percentage captured by Orange only
    real <- covid$Total.cases[j] / sum(covid$Total.cases)

    # sampled
    share <- c()
    for (i in 1:n) {
        share[i] <- ns[i,j] / sum(ns[i,])
    }
    
    # plot
    share %>% hist(ylab = "", main = covid$County[j])
    abline(v=real, col="red")
}
```

```{r}
par(mfrow = c(3,4))
n <- nrow(ns)

for (j in 25:36) {
    # percentage captured by Orange only
    real <- covid$Total.cases[j] / sum(covid$Total.cases)

    # sampled
    share <- c()
    for (i in 1:n) {
        share[i] <- ns[i,j] / sum(ns[i,])
    }
    
    # plot
    share %>% hist(ylab = "", main = covid$County[j])
    abline(v=real, col="red")
}
```

```{r}
par(mfrow = c(3,4))
n <- nrow(ns)

for (j in 37:48) {
    # percentage captured by Orange only
    real <- covid$Total.cases[j] / sum(covid$Total.cases)

    # sampled
    share <- c()
    for (i in 1:n) {
        share[i] <- ns[i,j] / sum(ns[i,])
    }
    
    # plot
    share %>% hist(ylab = "", main = covid$County[j])
    abline(v=real, col="red")
}
```

```{r}
par(mfrow = c(3,4))
n <- nrow(ns)

for (j in 49:58) {
    # percentage captured by Orange only
    real <- covid$Total.cases[j] / sum(covid$Total.cases)

    # sampled
    share <- c()
    for (i in 1:n) {
        share[i] <- ns[i,j] / sum(ns[i,])
    }
    
    # plot
    share %>% hist(ylab = "", main = covid$County[j])
    abline(v=real, col="red")
}
```

```{r}
#par(mfrow = c(3,3))
n <- nrow(ns)
probs_left <- c()
probs_right <- c()
count <- 1

for (j in c(1, 2, 5, 7, 10, 15, 18, 24, 26)) {
    # percentage captured by Orange only
    real <- covid$Total.cases[j] / sum(covid$Total.cases)

    # sampled
    share <- c()
    for (i in 1:n) {
        share[i] <- ns[i,j] / sum(ns[i,])
    }
    
    probs_left[count] <- sum(real < share) / n
    probs_right[count] <- sum(real > share) / n
    
    count <- count + 1
    # plot
    #share %>% hist(ylab = "", xlab = 'T(n)',main = covid$County[j])
    #abline(v=real, col="red")
}
```

```{r}
probs_left
probs_right
```




















